Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-12 popl24/automation.tex
checkpoint
Update automation.tex

Our first step in automating security analysis is to show how to
compute the basic program distribution. Recall that basic
distributions are based on the memories resulting from execution of an
$\minicat$ program, which contain a record of all client views and the
public output.  Given $\prog$ we denote this as $\runs(\prog)$, and we
denote the basic distribution of a program as $\progtt(\prog)$.
\begin{definition}
  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
  $$
  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
  $$
  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
  $\prog$, where given $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$
  we have for all $\store \in \mems(S \cup V \cup F)$:
  $$
  \progtt(\prog)(\store) =  1 / 2^{|S\cup F|} \ \text{if}\ \store \in \runs(\prog), \text{otherwise}\ 0
  $$
\end{definition}

@@

Since $|\runs(\prog)|$ grows exponentially in the size of
$\flips(\prog)$, scalability to larger programs is a concern.  However
computation of $\runs$ is amenable to optimizations, especially
parallelization- rather than computing the truth table
``horizontally'', it can be computed ``vertically''. We can rewrite
$\minifed$ programs $\prog$ to stratified Datalog programs, where
$\runs(\prog)$ is equivalent to set of least Herbrand models obtained
by representing different random tapes and input secrets as different
fact bases. Details of this rewrite are provided in Appendix
\ref{section-bruteforce}. The model for each fact base can be computed
independently and in parallel, with the potential to apply other
matrix optimizations developed in recent work on HPC for Datalog
\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.

\subsection{Certification of Extensional Properties}
\label{section-automation-extensional}

We will use \emph{certification} to refer to algorithmic methods of
verifying properties of $\minifed$ protocols, vs.~manual methods. For
the sake of clarity we make the following definition:
\begin{definition}
  A predicate $\phi$ on $\minifed$ protocols is \emph{certifiable} iff
  there exists an algorithm $A$ such that $A(\prog)$ iff $\phi(\prog)$,
  and a safe protocol $\prog$ is \emph{certified for $\phi$} iff
  $A(\prog)$ holds.
\end{definition}

>>>>

Our first step in automating security analysis is to show how to
compute the basic program distribution. Recall that basic
distributions are based on the memories resulting from the execution of an
$\minicat$ program, which contain a record of all client views and the
public output.  Given $\prog$ we denote this as $\runs(\prog)$, and we
denote the basic distribution of a program as $\progtt(\prog)$.
\begin{definition}
  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
  $$
  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
  $$
  We write $\progtt(\prog)$ to denote the \emph{basic distribution} of
  $\prog$, where given $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$
  we have for all $\store \in \mems(S \cup V \cup F)$:
  $$
  \progtt(\prog)(\store) =  1 / 2^{|S\cup F|} \ \text{if}\ \store \in \runs(\prog), \text{otherwise}\ 0
  $$
\end{definition}

@@

Since $|\runs(\prog)|$ grows exponentially in the size of
$\flips(\prog)$, scalability to larger programs is a concern.  However
computation of $\runs$ is amenable to optimizations, especially
parallelization- rather than computing the truth table
``horizontally'', it can be computed ``vertically''. We can rewrite
$\minifed$ programs $\prog$ to stratified Datalog programs, where
$\runs(\prog)$ is equivalent to the set of least Herbrand models obtained
by representing different random tapes and input secrets as different
fact bases. Details of this rewrite are provided in Appendix
\ref{section-bruteforce}. The model for each fact base can be computed
independently and in parallel, with the potential to apply other
matrix optimizations developed in recent work on HPC for Datalog
\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.

\subsection{Certification of Whole-Program Security Properties}
\label{section-automation-extensional}

We will use \emph{certification} to refer to algorithmic methods of
verifying properties of $\minifed$ protocols, vs.~manual methods. For
the sake of clarity, we make the following definition:
\begin{definition}
  A predicate $\phi$ on $\minifed$ protocols is \emph{certifiable} iff
  there exists an algorithm $A$ such that $A(\prog)$ iff $\phi(\prog)$,
  and a safe protocol $\prog$ is \emph{certified for $\phi$} iff
  $A(\prog)$ holds.
\end{definition}
